---
title: "Untitled"
author: "Montana"
date: "2024-09-09"
output: html_document
---

# Introduction 

This is an introduction to the components and the generation of the posterior distributions for the foodweb landscapes of Little Moose. Below you will find snippits of code and examples of the data frames that are used during this process. Here we will look at the data frames that are associated with the `step1_LML_source_L.R` file. However, the graphs will display the results of the posterior generation for both the early and late periods of this experiment.

--------------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F, include = F, warnings = F, message=F}
## Setup for the markdown
# Libraries ----------------------
library(easypackages)
libraries("snow","plotrix", "SIBER","ellipse","mixtools",
          "mvtnorm","plot3D","scatterplot3d","scales","viridis","ggplot2",
          "gridExtra", "MASS","plotly", "knitr","tidyverse", "mvtnorm","rjags")

```


```{r, echo = F, message=F, warning=F}

## Load in data -----------------


# Sources
# For functions
source("Data/Models/functions.R")
# For starting conditions. This tutorial is for the late period.
source("Landscapes/LateLandscapes/step1_LML_source_L.R")

## Posterior data frame --------------------
## Combining chains 

load("Data/LateData/LateJagsOutput.RData")

## Chains generated from previous script 
chains <- as.mcmc.list(jags_output.late) ## set chains
chains_combined <- gtable_combine(chains) ## combine chains

```

## Define starting conditions for the model snippits.

For all periods of this experiment, we will be looking at 20 posterior draws `n.posts`, 10 species `n_species`, 32 sites around the shore of Little Moose Lake `n_sites`, and 4 years `n_years`. The early period includes the years 2001-2004 and the late period includes the years 2019-2022.

```{r}
# Model conditions
n.posts = 20
n_species = 10
n_sites = 32
n_years = year_max - year_min -1 # seems missing 2023 data?

```

We begin by setting names for the data frame that holds the posterior distributions. These names are composed of useful information that we can later filter by. For example, `N_1_1_1` is the posterior distribution for the estimated true abundance of species 1 at site 1 during the first year of that period. 
```{r, echo = F, message=F, warning=F}
# Define names for the 
names = (data.frame(colnames = colnames(jags_output.late[[1]]))  %>%
           mutate(colnames = str_replace_all(colnames, "\\[", "_"), 
                  colnames = str_replace_all(colnames, "\\]", ""),
                  colnames = str_replace_all(colnames,",", "_")))$colnames

chain_dat = chains_combined %>%
  as.matrix() %>%
  as.data.frame() %>%
  set_names(names)


head(chain_dat)[1:5]

```
But the entire posterior distribution is a large data frame. At this point, we are not subsetting the posterior distribution and the abundances are estimated at each site, at each year, for each species. This results in a large number of columns.

```{r}
dim(chain_dat)
```

## Subsetting the compontents of the landscape generation.

We begin with sigma. Sigma represents the shape of the isotope ellipse for each species.It is composed of a covariance matrix.

Below we see an example of the sigma data from the posterior distribution. 

```{r, echo = F, messages = F, warning = F}
## Clean sigma estimates
sig.dat = chain_dat %>%
  select(contains("sigma", ignore.case = TRUE))  %>%
  as.data.frame() %>% 
  mutate(post = 1:length(.[,1])) %>%
  filter(post <= n.posts) %>%
  select(post, everything()) %>%
  pivot_longer(2:length(.[1,]), names_to = "metric", 
               values_to = "sig.value") %>%
  separate(metric,
           into = c("metric", "species", 
                    "sig.order1", "sig.order2")) %>% 
  mutate(metric = str_remove_all(metric, "[0-9]")) %>%
  unite("names", c(metric, sig.order1, sig.order2)) %>% 
  pivot_wider(names_from = names, values_from = sig.value)

head(sig.dat)
```
For `species == 1` and `post == 1`, see the below matrix as an example of what the covariance matrix looks like.

```{r, echo = F, message=F, warning = F}
cov.mat = sig.dat %>% 
  filter(post == 1, species ==1)

matrix(c(cov.mat$Sigma_1_1, cov.mat$Sigma_1_2, cov.mat$Sigma_2_1, cov.mat$Sigma_2_2),ncol = 2, nrow = 2)
```


The mass data is generated through the use of length and weight relationships for species in Little Moose Lake. The one exception here is brown bullhead. For brown bullhead, we use weight-length relationships that are available on FishBase from previously published research.

```{r}
#mass weight graph
```

See below the structure of the mass data.

```{r, echo = F, messages = F, warning = F}
## Clean length estimates
mass.dat = chain_dat %>%
  select(contains("avg_mass", ignore.case = TRUE))  %>%
  as.data.frame() %>%
  mutate(post = 1:length(.[,1])) %>%
  filter(post <= n.posts) %>%
  select(post, everything()) %>%
  pivot_longer(2:length(.[1,]),
               names_to = "metric", values_to = "mass.avg") %>%
  separate(metric, into = c("metric1","metric", "species"), sep = "_") %>%
  select(-metric1, -metric)

head(mass.dat)
```
Comparing the masses of fishes bewteen the two periods. Data from the posterior for both the late and the early period is loaded in here. 

```{r, message = F, echo = F, warning = F}
## Let's compare the early and late periods...
load("Data/posterior_early.csv")
load("Data/posterior_late.csv")
## Legend
species.late = c("BB", "CC","CS","LT","MM","PS","RS","SMB","SS","WS")
legend = data.frame(group = c(1:10), 
                    species  = species.late)

## Comparison of posterior parameter distributions

posterior.early$community = 1
posterior.late$community = 2

posterior = rbind(posterior.early, posterior.late)

posterior %>% 
  mutate(group = as.numeric(species)) %>%
  select(-species) %>%
  left_join(legend) %>%
  ggplot(aes(x = as.factor(community), y = mass.avg)) + 
  geom_boxplot() +
  scale_y_log10() +
  facet_wrap(~species, scales = "free") +
  xlab("Period") +
  ylab("Total Abundance") +
  theme_minimal(base_size = 12) +
  scale_x_discrete(labels = c("1" = "Early", "2" = "Late")) 

```

See below for the snippit of the abundance data.
```{r, echo = F, messages = F, warning = F}
# Clean abund estimates -- this does it year by year
abund.dat =  chain_dat %>%
  as.data.frame() %>%
  select(contains("N", ignore.case = TRUE)) %>%
  
  mutate(post = 1:length(.[,1])) %>%
  filter(post <= n.posts) %>%
  select(post, everything()) %>% 
  #
  pivot_longer(2:((n_species * n_years * n_sites)+1),
               names_to = "metric", 
               values_to = "value") %>%
  separate(metric, into = c("metric", "site", "year", "species")) %>%
  mutate(species = as.numeric(species),
         site = as.numeric(site),
         post = as.numeric(post), 
         year = as.numeric(year)) %>%
  group_by(post, species, year) %>%
  summarize(tot_abund = sum(value)) %>%
  mutate(species = as.character(species))

head(abund.dat)


```

See below for a comparison between the estimated total abundance for both periods.
```{r, echo = F, messages = F, warning = F}


posterior %>% 
  mutate(group = as.numeric(species)) %>%
  select(-species) %>%
  left_join(legend) %>%
  ggplot(aes(x = as.factor(community), y = tot_abund)) + 
  geom_boxplot() +
  scale_y_log10() +
  facet_wrap(~species, scales = "free") +
  xlab("Period") +
  ylab("Total Abundance") +
  theme_minimal(base_size = 12) +
  scale_x_discrete(labels = c("1" = "Early", "2" = "Late")) 

```

```{r, echo = F, messages = F, warning = F}
# Clean mu estimates
mu.dat = chain_dat %>%
  as.data.frame() %>% 
  select(contains("mu", ignore.case = TRUE)) %>%
  mutate(post = 1:length(.[,1])) %>%
  filter(post <= n.posts) %>%
  select(post, everything()) %>%
  pivot_longer(2:length(.[1,]), names_to = "metric", values_to = "mu.value") %>% 
  separate(metric, into = c("metric", "species", "isotope")) %>% 
  unite("names", c(metric, isotope)) %>%
  pivot_wider(names_from = names, values_from = mu.value)

head(mu.dat)

```




Finally, all the previous steps are combined. When applicable (i.e. where abundances are estimated across multiple years) the estimates are summarized by mean.

See below for the structure of the final posterior data frame is.
```{r, echo = FALSE, messages = FALSE, warning = FALSE, include = F}


## Final posterior frame
posterior.late = left_join(sig.dat, mass.dat) %>% 
  left_join(abund.dat) %>%
  left_join(mu.dat) %>%
  group_by(post, species) %>%
  select(year, everything()) %>%
  summarize(across(2:9, mean, na.rm = TRUE))



```
```{r, echo = FALSE, warning = FALSE, message=FALSE}
posterior.late %>% select(community, everything()) %>%
  head() 
```






